---
title: "repijson"
author: "Thomas Finnie; Andy South; Ellie Sherrard-Smith; Ana Bento, Thibaut Jombart"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

*EpiJSON* is a generic JSON format for storing epidemiological data.   

*repijson* is an R package that allows conversion between EpiJSON files and R data formats.

This vignette is a demonstration of the package *repijson*.


# Installing *repijson*
-------------
To install the development version from github:
```{r install, eval=FALSE}
library(devtools)
install_github("hackout2/repijson")
```


Then, to load the package, use:
```{r load}
library("repijson")
```

# The *EpiJSON* format

This is a simplified representation of the *EpiJSON* format.   

```{r, echo=FALSE, fig.width=7, fig.height=5, fig.path="figs/"}
#here deliberately not echoed just to show the diagram
epijsonObjectVis(textSize = 5) #textSize 4 is default
```

The *repijson* objects used to store *EpiJSON* are represented in the following diagram.
```{r, echo=FALSE, fig.width=5, fig.height=5, fig.path="figs/"}
#here deliberately not echoed just to show the diagram
epijsonObjectVis( attribMeta = 'ejAttribute',
                  attribRecord = 'ejAttribute',
                  attribEvent = 'ejAttribute',
                  labelObject = 'ejObject',
                  labelMeta = 'ejMetadata',
                  labelRecord = 'ejRecord',
                  labelEvent = 'ejEvent',
                  textSize = 5) #textSize 4 is default
```



# Validation

Demonstrating how we can go from a JSON character string or object to data frame to an obkData object

Load the required packages as necessary
```{r}

library(OutbreakTools)
library(sp)
library(HistData)
library(repijson)
```


These are example data in data.frame format

#Example: data frame 1
```{r}
data(Snow.deaths)
```
Adding some dates, pumps, some genders 
```{r}
simulated <- Snow.deaths
simulated$gender <- c("male","female")[(runif(nrow(simulated))>0.5) +1]
simulated$date <- as.POSIXct("1854-04-05") + rnorm(nrow(simulated), 10) * 86400
simulated$pump <- ceiling(runif(nrow(simulated)) * 5)

exampledata1<-head(simulated)
exampledata1
```

#Example: data.frame 2
```{r}
exampledata2<- data.frame(id=c("A","B","3D","4d"),
                 name=c("tom","andy","ellie","Ana"),
                 dob=c("1984-03-14","1985-11-13","1987-06-16","1987-06-16"),
                 gender=c("male","male","female","female"),
                 rec1contact=c(2,1,5,1),
                 rec1date=c("2014-12-28","2014-12-29","2015-01-03","2015-01-08"),
                 rec1risk=c("high","high","low","high"),  
                 rec1temp=c(39,41,41,39),
                 rec2contact=c(4,1,1,1),
                 rec2date=c("2015-01-02","2015-01-12","2015-01-09","2015-01-09"),
                 rec2risk=c("high","low","high","low"),stringsAsFactors=FALSE)
exampledata2
```

################################################
## Transition 1: data.frame to EpiJSON format ##
################################################

Use the *repijson* package to convert a data.frame object into a EpiJSON object within R:

```{r}	
eg1 <- as.ejObject(exampledata1,	
    recordAttributes = "gender",	
    eventDefinitions = list(define_ejEvent(date="date",	name=NA, location=list(x="x", y="y", proj4string=""), attributes="pump")),
 		metadata=list())		       
eg1
```



```{r}
eg2 <- as.ejObject(exampledata2, recordAttributes = c("name","dob","gender"),
     eventDefinitions = list(define_ejEvent(name="rec1contact",date="rec1date", attributes=list("rec1risk","rec1temp")),
                             define_ejEvent(name="rec2contact",date="rec2date", attributes="rec2risk")),
 		metadata=list())
eg2
```


#######################################################
## Transition 2: EpiJSON object to data.frame format ##
#######################################################

Use the *repijson* package to convert a JSON object into a data.frame object:
```{r}
as.data.frame(eg1)
```

#######################################################
## Transition 3: From obkData to an EpiJSON object   ##
#######################################################

These are example data in obkData format
```{r}
data(ToyOutbreak) 
```

Use the *repijson* package to convert an obkData object to JSON object into :
```{r}
eg3 <- as.ejObject(ToyOutbreak)
```

#######################################################
## Transition 4: From an EpiJSON object to obkData   ##
#######################################################

*Next function to produce*


#######################################################
## Transition 5: From an EpiJSON object to spatial   ##
#######################################################

Use the *repijson* package to convert from an EpiJSON object to spatial (sp)
```{r}
example1 <- as.ejObject(simulated, recordAttributes = c("gender"),
     eventDefinitions = list(define_ejEvent(date="date", name=NA, location=list(x="x", y="y", proj4string=""), attributes="pump")),
 		metadata=list())

#this is causing failure with
#Error in as.data.frame.default(X[[1L]], ...) : 
#  cannot coerce class ""ejAttribute"" to a data.frame
#sp_eg1 <- as.SpatialPointsDataFrame.ejObject(example1)

#plot(sp_eg1,pch=20,col="green")
#text(10,17,"Example from Snow Deaths data")

```

*The story continues...!!!*