---
title: "repijson"
author: "Thomas Finnie; Andy South; Ellie Sherrard-Smith; Ana Bento, Thibaut Jombart"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

*EpiJSON* is a generic JSON format for storing epidemiological data.   

*repijson* is an R package that allows conversion between EpiJSON files and R data formats.

This vignette is a demonstration of the package *repijson*.


# Installing *repijson*
-------------
To install the development version from github:
```{r install, eval=FALSE}
library(devtools)
install_github("hackout2/repijson")
```


Then, to load the package, use:
```{r load}
library("repijson")
```

# The *EpiJSON* format

This is a simplified representation of the *EpiJSON* format.   

```{r, echo=FALSE, fig.width=7, fig.height=5, fig.path="figs/"}
#here deliberately not echoed just to show the diagram
epijsonObjectVis(textSize = 10) #textSize 4 is default
```

The *repijson* objects used to store *EpiJSON* are represented in the following diagram.
```{r, echo=FALSE, fig.width=5, fig.height=5, fig.path="figs/"}
#here deliberately not echoed just to show the diagram
epijsonObjectVis( attribMeta = 'ejAttribute',
                  attribRecord = 'ejAttribute',
                  attribEvent = 'ejAttribute',
                  labelObject = 'ejObject',
                  labelMeta = 'ejMetadata',
                  labelRecord = 'ejRecord',
                  labelEvent = 'ejEvent',
                  textSize = 5) #textSize 4 is default
```



# Validation

Demonstrating how we can go from a JSON character string or object to data frame to an obkData object

Load the required packages as necessary
```{r}

library(OutbreakTools)
library(sp)
library(HistData)
library(repijson)
```


These are example data in data.frame format

#Example: data frame 1
```{r}
data(Snow.deaths)
```
Adding some dates, pumps, some genders 
```{r}
simulated <- Snow.deaths
simulated$gender <- c("male","female")[(runif(nrow(simulated))>0.5) +1]
simulated$date <- as.POSIXct("1854-04-05") + rnorm(nrow(simulated), 10) * 86400
simulated$pump <- ceiling(runif(nrow(simulated)) * 5)

exampledata1<-head(simulated)
exampledata1
```

#Example: data.frame 2
```{r}
exampledata2<- data.frame(id=c(1,2,3,4,5),
                 name=c("Tom","Andy","Ellie","Ana","Tibo"),
                 dob=c("1981-01-12","1980-11-11","1982-02-10","1981-12-09","1983-03-08"),
                 gender=c("male","male","female","female","male"),
                 date.of.onset=c("2014-12-28","2014-12-29","2015-01-03","2015-01-08","2015-01-04"),
                 date.of.admission=c(NA,"2015-01-05","2015-01-12",NA,"2015-01-14"),
                 date.of.discharge=c(NA,NA,"2015-01-17",NA,"2015-01-17"),
                 hospital=c(NA,"St Marys","Whittington",NA,"Whittington"),
                 fever=c("yes","yes","no","no","yes"),
                 sleepy=c("no","yes","yes","no","yes"),
                 contact1.id=c("B","A","5",NA,"3D"),
                 contact1.date=c("2014-12-26","2014-12-26","2014-12-28",NA,"2014-12-28"),
                 contact2.id=c("3D","3D","5",NA,"A"),
                 contact2.date=c("2014-12-25","2014-12-26","2015-01-14",NA,"2014-12-25"),
                 contact3.id=c("B",NA,NA,NA,NA),
                 contact3.date=c("2015-01-08",NA,NA,NA,NA)
                 )
exampledata2
```

################################################
## Transition 1: data.frame to EpiJSON format ##
################################################

Use the *repijson* package to convert a data.frame object into a EpiJSON object within R:

```{r}	
eg1 <- as.ejObject(exampledata1,	
    recordAttributes = "gender",	
    eventDefinitions = list(define_ejEvent(id="case",name=NA, location=list(x="x", y="y", proj4string=""), attributes="pump")),
 		metadata=list())		       
eg1
```



```{r}
eg2 <- as.ejObject(exampledata2, recordAttributes = c("id","name","dob","gender"),
     eventDefinitions = list(define_ejEvent(name="date.of.onset",date="date.of.onset", 
                                            attributes=list("hospital","fever","sleepy")),
                             define_ejEvent(name="contact1.id",date="contact1.date", attributes="")),
 		metadata=list())
eg2
```


#######################################################
## Transition 2: EpiJSON object to data.frame format ##
#######################################################

Use the *repijson* package to convert a JSON object into a data.frame object:
```{r}
as.data.frame(eg1)
```

#######################################################
## Transition 3: From obkData to an EpiJSON object   ##
#######################################################

These are example data in obkData format
```{r}
data(ToyOutbreak) 
```

Use the *repijson* package to convert an obkData object to JSON object into :
```{r}
eg3 <- as.ejObject(ToyOutbreak)
```

#######################################################
## Transition 4: From an EpiJSON object to obkData   ##
#######################################################

*Next function to produce*


#######################################################
## Transition 5: From an EpiJSON object to spatial   ##
#######################################################

Use the *repijson* package to convert from an EpiJSON object to spatial (sp)
```{r}
example1 <- as.ejObject(simulated, recordAttributes = c("gender"),
     eventDefinitions = list(define_ejEvent(date="date", name=NA, location=list(x="x", y="y", proj4string=""), attributes="pump")),
 		metadata=list())

#this is causing failure with
#Error in as.data.frame.default(X[[1L]], ...) : 
#  cannot coerce class ""ejAttribute"" to a data.frame
#sp_eg1 <- as.SpatialPointsDataFrame.ejObject(example1)

#plot(sp_eg1,pch=20,col="green")
#text(10,17,"Example from Snow Deaths data")

```

*The story continues...!!!*